---
title: "BTE Appendix I"
author: "Ze-Yi Han"
date: "6/14/2022"
output:
  html_document:
    fig_caption: yes
    highlight: pygments
    theme: spacelab
    toc: yes
    toc_depth: 3
    code_folding: show
---

<style>
h1, h2, h3, h4, h5, h6 {
    font-family: "Bookman", "Gerogia";
    color: ##69b3a2
}

a, a:hover {
    color: ##b24e00;
}

pre {
    font-size: 12px;
    color: ##546399;
}
</style>


---
```{r setup, echo = FALSE}
### Loadings----
knitr::opts_chunk$set(eval = TRUE, cache = TRUE, warning=FALSE, message=FALSE,fig.width=6, fig.height=10, fig.align = "center")
pacman::p_load(ggplot2, psych, reshape2, cowplot,dplyr,tidyr,data.table,readr,mgcv,nlme,broom,rEDM, shape)
```

# Data 

Data read in

We used log density to standardize the variance in protists density time series. Since in some days, the protists density is 0, and will result -INF by log(), we added 1 to all the density before calculating the log density for model fitting. We minused the 1 after exponentiating the model predictions. 
```{r}
### data read in------
## density 
milou_dens0 <- read.csv("/Users/zeyihan/Documents/PhD/Projects/Milou/02_Clean_data/Milou_density.csv")
milou_dens <- milou_dens0  %>% mutate(tre=as.factor(tre), temp = as.factor(temp), nut = as.factor(nut), all_rep = rep(1:24, time = 24), log_density = log(density + 1)) %>% select(-1)  
tetra <- milou_dens[which(milou_dens$Class=="Tetrahymena"),]
eup <- milou_dens[which(milou_dens$Class=="Euplotes"),]
bac0 <- read.csv("/Users/zeyihan/Documents/PhD/Projects/Milou/02_Clean_data/Milou_bac_OD.csv")
bac1 <- bac0 %>% mutate(tre = as.factor(tre), temp = as.factor(temp), nut = as.factor(nut), all_rep = rep(1:24, time = 11), temp_num = as.numeric(rep(c(22,25), each = 6, times = 22)), nut_num = as.numeric(rep(c(1,0.5), each = 12, times = 11)))
```

We calculated the mean and standard variation for bacteria OD after day 5. One data point that is more than 3 standard deviation away from the mean is removed. 

```{r}
sd <- sd(bac1$OD[which(bac1$day > 5)], na.rm = TRUE)
mean <- mean(bac1$OD[which(bac1$day > 5)], na.rm = TRUE)
bac <- bac1[-which(bac1$OD > mean+3*sd),]
```

Since after population collapse, Tetrahymena are functionally extinct and so, even if we occasionally found individual, trait distributions are unreliable. For traits observation made after each Tetrahymena treatment collapsed (once population density became 0 ind/ml) that with less than 10 number of individual, we considered them unreliable and, thus, removed for future analysis. 

```{r }
## trait
t_summ0 <- fread("/Users/zeyihan/Documents/PhD/Projects/Milou/02_Clean_data/Milou_trait_summary.csv")
t_summ <- t_summ0 %>% mutate(rep = as.character(rep),
         tre = as.factor(tre),temp_num = as.numeric(temp_num), nut_num = as.numeric(nut_num),rep = as.factor(rep), tre = as.factor(tre), all_rep = rep(1:24, time = 24), temp = as.factor(temp), nut = as.factor(nut))
tetraSumm <- t_summ %>% filter(Class =="Tetrahymena") 
eupSummtrait <- t_summ %>% filter(Class =="Euplotes")

# replace NA for tetra trait}
# Half 22C
tetraSumm[, 14:20][which(tetraSumm$tre=="Half22C" & tetraSumm$day>10 & tetraSumm$num_ind<10)] <- NA
# Full 22C
tetraSumm[, 14:20][which(tetraSumm$tre=="Full22C" & tetraSumm$day>11 & tetraSumm$num_ind<10)] <- NA
# Half 25C
tetraSumm[, 14:20][which(tetraSumm$tre=="Half25C" & tetraSumm$day>8 & tetraSumm$num_ind<10)] <- NA
# Full 25C
tetraSumm[, 14:20][which(tetraSumm$tre=="Full25C" & tetraSumm$day>8 & tetraSumm$num_ind<10)] <- NA

# trait group mean
tetraSummMean <- tetraSumm %>% na.omit %>% group_by(Class,temp,nut,tre,day) %>% summarise(mArea = mean(area)) %>% filter(day <= 9)
eupSummMean <- eupSummtrait %>% na.omit %>% group_by(Class,temp,nut,tre,day) %>% summarise(mArea = mean(area)) %>% filter(day <= 9) %>% filter(day <= 9)

# cropped time series for second version of CCM
tetraSumm_crop <- tetraSumm %>% filter(day <= 9)
eupSumm_crop <- eupSummtrait %>% filter(day <=9)
```


Calculate mean density of Bacteria, Tetrahymena, and Euplotes
```{r}
bac.2 <- bac %>%
  rename(density = OD) %>% 
  select(tre:density,treatNum) %>%
  mutate(temp = as.factor(temp))

teb.all <- milou_dens %>% 
 select(tre:density,temp_num,-temp,treatNum)  %>%
  rename(temp = temp_num) %>%
  mutate(temp = as.factor(temp), nut = as.factor(nut),day = as.numeric(day))%>%
  full_join(bac.2) %>% 
  na.omit() %>%
  group_by(Class,tre,treatNum, nut, temp,day) %>%
  summarise(mean.d = mean(density),
            log.d = log(mean.d +1)) %>%
  mutate(day = as.numeric(day))

teb.mean <- teb.all %>%
 select(Class:mean.d) %>%
  spread(key=Class, value = mean.d)
```

# 1. Appendix I Table S1. Gamm models for Density Data

We used GAMM model to select for the detection of temporal autocorrelation and correctly accounting for that. To do so, we treated replicates as random effects and day and treatment as fixed effects. Temporal autocorrelation was accounted for using Autoregressive Moving Average (ARMA) correlation structure (of orders 1, 2 and 3) using “corARMA” in R package “nlme” to apply in the GAMM models using the ‘nlme’ R package (version 3.1-148). GAMM models smoothed by treatment more closely reproduce the observed dynamics and are thus better suited for the detection autocorrelation in the time series data, as well as for time series analysis using CCM.

## 1.1 Bacteria Density ARMA selection

AR = 2 is the most suitable for Bacteria Density. 

```{r Bac gamm}
#bac1 <- bac1 %>% mutate(tre = as.factor(tre), temp_num = as.numeric(temp), nut_num = as.numeric(rep(c(1,0.5), each = 12, time = 11 )))
bm0 <- gamm(OD ~ s(day, by = tre, k =10) + tre ,random = list(rep= ~ 1), data = bac) 
summary(bm0$gam)

bm1 <- gamm(OD ~ s(day, by = tre, k =10) + tre ,random = list(rep= ~ 1), data = bac, correlation = corARMA(form = ~ day|tre/rep, p = 1)) 
summary(bm1$gam)

#AR2  
bm2 <- gamm(OD ~ s(day, by = tre, k =10) + tre,random = list(rep= ~ 1), data = bac, correlation = corARMA(form = ~ day|tre/rep, p = 2)) 
summary(bm2$gam)

#AR3
bm3 <- gamm(OD ~ s(day, by = tre, k =10) + tre ,random = list(rep= ~ 1), data = bac, correlation = corARMA(form = ~ day|tre/rep, p = 3)) 
summary(bm3$gam)

# Comparing model result
AIC(bm1$lme,bm2$lme,bm3$lme, bm0$lme) 
```

## 1.2 Tetrahymena Density ARMA selection

We tested gamm models on Tetrahymena as we tested on bacteria. AR = 1 has the lowest AIC. 
```{r tetra AR(0)}
# no autocorrelation accounted for
tm0 <- gamm(log_density ~ s(day, by = tre, k =10) + tre,random = list(rep= ~ 1), data = tetra) 
summary(tm0$gam)

#AR1
tm1 <- gamm(log_density ~ s(day, by = tre, k =10) + tre ,random = list(rep= ~ 1), data = tetra, correlation = corARMA(form = ~ day|tre/rep, p = 1)) 
summary(tm1$gam)

#AR2  
tm2 <- gamm(log_density ~ s(day, by = tre, k =10) + tre,random = list(rep= ~ 1), data = tetra, correlation = corARMA(form = ~ day|tre/rep, p = 2)) 
summary(tm2$gam)

#AR3
tm3 <- gamm(log_density ~ s(day, by = tre, k =10) + tre ,random = list(rep= ~ 1), data = tetra, correlation = corARMA(form = ~ day|tre/rep, p = 3)) 
summary(tm3$gam)

# Comparing model result
AIC(tm0$lme,tm1$lme,tm2$lme,tm3$lme )
```


## 1.3  Euplotes Density ARMA selection

Same models as above. Gamm model without AR structure has the lowest AIC.
```{r eup AR(0)}
eup$log_density <- log(eup$density + 1)
em0 <- gamm(log_density ~ s(day, by = tre, k =10) + tre,random = list(rep= ~ 1), data = eup) 
summary(em0$gam)

#AR1
em1 <- gamm(log_density ~ s(day, by = tre, k =10) + tre ,random = list(rep= ~ 1), data = eup, correlation = corARMA(form = ~ day|tre/rep, p = 1)) 
summary(em1$gam)

#AR2  
em2 <- gamm(log_density ~ s(day, by = tre, k =10) + tre,random = list(rep= ~ 1), data = eup, correlation = corARMA(form = ~ day|tre/rep, p = 2)) 
summary(em2$gam)

#AR3
em3 <- gamm(log_density ~ s(day, by = tre, k =10) + tre ,random = list(rep= ~ 1), data = eup, correlation = corARMA(form = ~ day|tre/rep, p = 3)) 
summary(em3$gam)

# Comparing model result
anova(em0$lme,em1$lme,em2$lme,em3$lme)
```

## 1.4 Tetrahymena Area ARMA selection

Same models as above but for Tetrahymena phenotypic data. Gamm model with AR = 3 structure has the lowest AIC.
```{r tetra area gamm model 1}
# model with interactive effect from 4 treatments
# No AR
tt0 <- gamm(area ~ s(day, by = tre, k =7) + tre ,random = list(rep= ~ 1), data = tetraSumm) 
summary(tt0$gam)
# AR = 1
tt1 <- gamm(area ~ s(day, by = tre, k =7) + tre,random = list(rep= ~ 1), data = tetraSumm, correlation = corARMA(form = ~ day|tre/rep, p = 1))
summary(tt1$gam)
# AR = 2
tt2 <- gamm(area ~ s(day, by = tre, k =7) + tre,random = list(rep= ~ 1), data = tetraSumm, correlation = corARMA(form = ~ day|tre/rep, p = 2))
summary(tt2$gam)
# AR = 3
tt3<- gamm(area ~ s(day, by = tre, k =7) + tre,random = list(rep= ~ 1), data = tetraSumm, correlation = corARMA(form = ~ day|tre/rep, p = 3))
summary(tt3$gam)
# Model comparison
AIC(tt0$lme, tt1$lme, tt2$lme, tt3$lme) # pick AR(3) 
```


## 1.5 Euplotes Area ARMA Selection

Same models as above but for Euplotes phenotypic data. Gamm model with AR = 1 structure has the lowest AIC.

```{r eup area gamm, fig.width = 3,fig.height=2}
# interactive effects from different treatment
# No AR
et0 <- gamm(area ~ s(day, by=tre, k =9) + tre,random = list(rep= ~ 1), data = eupSummtrait) 
summary(et0$gam)
# AR = 1
et1 <- gamm(area ~ s(day, by=tre, k =9) + tre,random = list(rep= ~ 1), data = eupSummtrait, correlation = corARMA(form = ~ day|tre/rep, p = 1)) 
summary(et1$gam)
# AR = 2
et2 <- gamm(area ~ s(day, by=tre, k =9) + tre,random = list(rep= ~ 1), data = eupSummtrait, correlation = corARMA(form = ~ day|tre/rep, p = 2)) 
summary(et2$gam)
# AR = 3
et3 <- gamm(area ~ s(day, by=tre, k =9) + tre,random = list(rep= ~ 1), data = eupSummtrait, correlation = corARMA(form = ~ day|tre/rep, p = 3)) 
summary(et3$gam)

# model compariton
AIC(et0$lme,et1$lme,et2$lme,et3$lme ) # pick AR 1
```
# 2. Appendix I Table S2. Quantifying temperature and nutrient effects on species temporal ecological and phenotypic dynamics using ARMA GAMM models. 
 
Because we have only two temperature and nutrients levels, we treated them as categorical variables. Here we present model comparison using AICc on models that accounts for all combinations of additive and/or interactive effects temperature and nutrient. We apply the ARMA structure selected in these models. 

## 2.1 Bacteria Density

Model 7 with temperature additive and nutrient interactive effects (nutrients within smooth term) is the most parsimonious model. 

```{r, results = "hide"}
# 1. Temp additive effect 
b.od1 <- gamm(OD ~ s(day, k = 10) + temp, random = list(all_rep = ~1), correlation = corARMA(form = ~ day|all_rep, p = 2),data = bac)
summary(b.od1$gam)

# 2. Nut additive effect 
b.od2 <- gamm(OD ~ s(day, k = 10) + nut,random = list(all_rep= ~ 1),correlation = corARMA(form = ~ day|all_rep, p = 2), data = bac)
summary(b.od2$gam)   # nutrient not significant

# 3. Temp and Nut additive effects 
b.od3 <- gamm(OD ~ s(day, k = 10) + nut + temp,random = list(all_rep= ~ 1), correlation = corARMA(form = ~ day|all_rep, p = 2),data = bac)
summary(b.od3$gam)  # nutrient not significant

# 4. Temp interactive effect ONLY (Temp within smooth term)
b.od4 <- gamm(OD ~ s(day, by =  temp, k = 10) + temp, random = list(all_rep = ~1), correlation = corARMA(form = ~ day|all_rep, p = 2),data = bac)
summary(b.od4$gam)

# 5. Nut interactive effect ONLY (Nut within smooth term)
b.od5 <- gamm(OD ~ s(day, by =  nut, k = 10) + nut , random = list(all_rep = ~1), correlation = corARMA(form = ~ day|all_rep, p = 2),data = bac)
summary(b.od5$gam) 

# 6. Temp interactive + nut additive (Temp within smooth term)
b.od6 <- gamm(OD ~ s(day, by =  temp, k = 10) + nut+ temp, random = list(all_rep = ~1),correlation = corARMA(form = ~ day|all_rep, p = 2), data = bac)
summary(b.od6$gam) 

# 7. Nut interactive + temp additive (Nut within smooth term)
b.od7 <- gamm(OD ~ s(day, by =  nut, k = 10) + temp+ nut, random = list(all_rep = ~1), correlation = corARMA(form = ~ day|all_rep, p = 2),data = bac)
summary(b.od7$gam) 

# 8. nut * temp outside of smooth term. 
b.od8 <- gamm(OD ~ s(day, k = 10) + temp*nut, random = list(all_rep = ~1), correlation = corARMA(form = ~ day|all_rep, p = 2),data = bac)
summary(b.od8$gam) 

# Model Comparison
AIC(b.od1$lme, b.od2$lme,b.od3$lme, b.od4$lme,b.od5$lme,b.od6$lme,b.od7$lme, b.od8$lme)
```

## 2.2 Tetrahymena Density

Model 6 with nutrient additive and temperature interactive effects (temperature within smooth term) is the most parsimonious model. 

```{r, results = "hide"}
# 1. Temp additive effect
t.dens1 <- gamm(log_density ~ s(day, k = 10) + temp, random = list(all_rep = ~1), data = tetra, correlation = corARMA(form = ~ day|temp/all_rep, p = 1))
summary(t.dens1$gam)

# 2. Nut additive effect 
t.dens2 <- gamm(log_density ~ s(day, k = 10) + nut,random = list(all_rep= ~ 1), data = tetra, correlation = corARMA(form = ~ day|nut/all_rep, p = 1))
summary(t.dens2$gam)   

# 3. Temp and Nut additive effects 
t.dens3 <- gamm(log_density ~ s(day, k = 10) + nut + temp,random = list(all_rep= ~ 1), data = tetra, correlation = corARMA(form = ~ day|all_rep, p = 1))
summary(t.dens3$gam)  

# 4. Temp interactive effect ONLY (Temp within smooth term)
t.dens4 <- gamm(log_density ~ s(day, by =  temp, k = 10) + temp , random = list(all_rep = ~1), data = tetra, correlation = corARMA(form = ~ day|temp/all_rep, p = 1))
summary(t.dens4$gam)

# 5. Nut interactive effect ONLY (Nut within smooth term)
t.dens5 <- gamm(log_density ~ s(day, by =  nut, k = 10) + nut, random = list(all_rep = ~1), data = tetra, correlation = corARMA(form = ~ day|nut/all_rep, p = 1))
summary(t.dens5$gam) 

# 6. Temp interactive + nut additive  (Temp within smooth term)
t.dens6 <- gamm(log_density ~ s(day, by =  temp, k = 10) + nut + temp, random = list(all_rep = ~1), data = tetra, correlation = corARMA(form = ~ day|all_rep, p = 1))
summary(t.dens6$gam) 

# 7. Nut interactive + temp additive (Nut within smooth term)
t.dens7 <- gamm(log_density ~ s(day, by =  nut, k = 10) + temp + nut, random = list(all_rep = ~1), data = tetra, correlation = corARMA(form = ~ day|all_rep, p = 1))
summary(t.dens7$gam) 

# 8. temp*nut outside of smooth term
t.dens8 <- gamm(log_density ~ s(day,  k = 10) + temp*nut, random = list(all_rep = ~1), data = tetra, correlation = corARMA(form = ~ day|all_rep, p = 1))
summary(t.dens8$gam) 

# model comparison
AIC(tm1$lme,t.dens1$lme, t.dens2$lme,t.dens3$lme, t.dens4$lme,t.dens5$lme, t.dens6$lme,t.dens7$lme,t.dens8$lme)
```
## 2.3  Euplotes Density

Model 4 with temperature interactive effects only (temperature within smooth term) is the most parsimonious model. 

```{r, results = "hide"}
# 1. Temp additive effect
e.dens1 <- gamm(log_density ~ s(day, k = 10) + temp, random = list(all_rep = ~1), data = eup)
summary(e.dens1$gam)

# 2. Nut additive effect 
e.dens2 <- gamm(log_density ~ s(day, k = 10) + nut,random = list(all_rep= ~ 1), data = eup)
summary(e.dens2$gam)   # nutrient not significant

# 3. Temp and Nut additive effects 
e.dens3 <- gamm(log_density ~ s(day, k = 10) + nut + temp,random = list(all_rep= ~ 1), data = eup)
summary(e.dens3$gam)  # nutrient not significant

# 4. Temp interactive effect ONLY (Temp within smooth term)
e.dens4 <- gamm(log_density ~ s(day, by =  temp, k = 10) + temp, random = list(all_rep = ~1), data = eup)
summary(e.dens4$gam)

# 5. Nut interactive effect ONLY (Nut within smooth term)
e.dens5 <- gamm(log_density ~ s(day, by =  nut, k = 10) + nut, random = list(all_rep = ~1), data = eup)
summary(e.dens5$gam) 

# 6. Temp interactive + nut additive (Temp within smooth term)
e.dens6 <- gamm(log_density ~ s(day, by =  temp, k = 10) + nut + temp, random = list(all_rep = ~1), data = eup)
summary(e.dens6$gam) # nutrient is not significant, temperature is

# 7. Nut interactive + temp additive (Nut within smooth term)
e.dens7 <- gamm(log_density ~ s(day, by =  nut, k = 10) + temp+ nut, random = list(all_rep = ~1), data = eup)
summary(e.dens7$gam) 

# 8. Nut * temp outside of smooth
e.dens8 <- gamm(log_density ~ s(day, k = 10) + temp*nut, random = list(all_rep = ~1), data = eup)
summary(e.dens8$gam) 

# Model Comparison
AIC(e.dens1$lme, e.dens2$lme,e.dens3$lme, e.dens4$lme,e.dens5$lme, e.dens6$lme,e.dens7$lme,e.dens8$lme)
```

## 2.4 Tetrahymena Area 

Same model as above, but for Tetrahymena phenotypic data.Model 7 with temperature additive and nutrient interactive effects (nutrients within smooth term) has the lowest AIC. 

```{r}
# 1. Temp additive effect
t.area1 <- gamm(area ~ s(day, k = 7) + temp, random = list(all_rep = ~1), data = tetraSumm,correlation = corARMA(form = ~ day|temp/all_rep, p = 3))
summary(t.area1$gam)
class(tetraSumm$temp)
# 2. Nut additive effect 
t.area2 <- gamm(area ~ s(day, k = 7) + nut,random = list(all_rep= ~ 1), data = tetraSumm, correlation = corARMA(form = ~ day|nut/all_rep, p = 3))
summary(t.area2$gam)   

# 3. Temp and Nut additive effects 
t.area3 <- gamm(area ~ s(day, k = 7) + nut + temp,random = list(all_rep= ~ 1), data = tetraSumm, correlation = corARMA(form = ~ day|all_rep, p = 3))
summary(t.area3$gam)  

# 4. Temp interactive effect ONLY
t.area4 <- gamm(area ~ s(day, by =  temp, k = 7) + temp, random = list(all_rep = ~1), data = tetraSumm, correlation = corARMA(form = ~ day|temp/all_rep, p = 3))
summary(t.area4$gam)

# 5. Nut interactive effect ONLY
t.area5 <- gamm(area ~ s(day, by =  nut, k = 7) + nut, random = list(all_rep = ~1), data = tetraSumm, correlation = corARMA(form = ~ day|nut/all_rep, p = 3))
summary(t.area5$gam) 

# 6. Temp interactive + nut additive
t.area6 <- gamm(area ~ s(day, by =  temp, k = 7) + nut + temp, random = list(all_rep = ~1), data = tetraSumm, correlation = corARMA(form = ~ day|all_rep, p = 3))
summary(t.area6$gam) 

# 7. Nut interactive + temp additive
t.area7 <- gamm(area ~ s(day, by =  nut, k = 7) + temp + nut, random = list(all_rep = ~1), data = tetraSumm, correlation = corARMA(form = ~ day|all_rep, p =3))
summary(t.area7$gam) 

# 8. Nut * temp 
t.area8 <- gamm(area ~ s(day,  k = 7) + temp * nut, random = list(all_rep = ~1), data = tetraSumm, correlation = corARMA(form = ~ day|all_rep, p = 3))
summary(t.area8$gam)

# model comparison
AIC(t.area1$lme, t.area2$lme,t.area3$lme, t.area4$lme,t.area5$lme, t.area6$lme,t.area7$lme, t.area8$lme)
```

## 2.5 Euplotes Area

Same model as above, but for Euplotes phenotypic data. Model 6 with nutrient additive and temperature interactive effects (temperature within smooth term) is the most parsimonious model. 
```{r}
# 1. Temp additive effect
eup.area1 <- gamm(area ~ s(day, k = 9) + temp, random = list(all_rep = ~1), data = eupSummtrait,correlation = corARMA(form = ~ day|temp/all_rep, p = 1))
summary(eup.area1$gam)

# 2. Nut additive effect 
eup.area2 <- gamm(area ~ s(day, k = 9) + nut,random = list(all_rep= ~ 1), data = eupSummtrait, correlation = corARMA(form = ~ day|nut/all_rep, p = 1))
summary(eup.area2$gam)   

# 3. Temp and Nut additive effects 
eup.area3 <- gamm(area ~ s(day, k = 9) + nut + temp,random = list(all_rep= ~ 1), data = eupSummtrait, correlation = corARMA(form = ~ day|all_rep, p = 1))
summary(eup.area3$gam)  

# 4. Temp interactive effect ONLY
eup.area4 <- gamm(area ~ s(day, by =  temp, k = 9) + temp, random = list(all_rep = ~1), data = eupSummtrait, correlation = corARMA(form = ~ day|temp/all_rep, p = 1))
summary(eup.area4$gam)

# 5. Nut interactive effect ONLY
eup.area5 <- gamm(area ~ s(day, by =  nut, k = 9) + nut, random = list(all_rep = ~1), data = eupSummtrait, correlation = corARMA(form = ~ day|nut/all_rep, p = 1))
summary(eup.area5$gam) 

# 6. Temp interactive + nut additive
eup.area6 <- gamm(area ~ s(day, by =  temp, k = 9) + nut + temp, random = list(all_rep = ~1), data = eupSummtrait, correlation = corARMA(form = ~ day|all_rep, p = 1))
summary(eup.area6$gam) 

# 7. Nut interactive + temp additive
eup.area7 <- gamm(area ~ s(day, by =  nut, k = 9) + temp + nut, random = list(all_rep = ~1), data = eupSummtrait, correlation = corARMA(form = ~ day|all_rep, p = 1))
summary(eup.area7$gam) 

# 8. nut * temp outside of smooth term
eup.area8 <- gamm(area ~ s(day, k = 9) + temp *nut, random = list(all_rep = ~1), data = eupSummtrait, correlation = corARMA(form = ~ day|all_rep, p = 1))
summary(eup.area8$gam) 

# model comparison
AIC(eup.area1$lme, eup.area2$lme,eup.area3$lme, eup.area4$lme,eup.area5$lme, eup.area6$lme,eup.area7$lme, eup.area8$lme)
```


# 3. Appendix I Table S3. Linear Models for Ecological Dynamics Descriptor

## 3.1 Bacteria OD600 

The following are ecological descriptor calculated for each species. 

```{r bacteria simp desc}
## New data frame
bac_desc <- data.frame("tre" = rep(c("Full22C", "Full25C", "Half22C", "Half25C"), each = 6))
bac_desc<- bac_desc %>%
  mutate(temp = as.factor(rep(c(22,25), each=6, times=2)), 
         nut = factor(rep(c("Full","Half"), each=12), levels = c("Half", "Full")), 
         rep = rep(1:6, times = 4))

## initial growth rate of Bacteria from day0 to day 1
bac_day0 <- bac[which(bac$day == 0), ]
bac_day1 <- bac[which(bac$day == 1), ]
bac_desc$init_growth <- (bac_day1$OD-bac_day0$OD)/8

## max abundance: three days with the top three population density
bac_max1<- teb.mean%>% group_by(tre,treatNum) %>% select(tre:day,Bacteria) %>% na.omit() %>%  arrange(desc(Bacteria), .by_group = TRUE)%>% slice_max(Bacteria, n=3) 

bm1 <- bac %>% filter(tre == "Full22C", day %in% bac_max1$day[which(bac_max1$tre=="Full22C")])
bm2 <- bac %>% filter(tre == "Full25C", day %in% bac_max1$day[which(bac_max1$tre=="Full25C")])
bm3 <- bac %>% filter(tre == "Half22C", day %in% bac_max1$day[which(bac_max1$tre=="Half22C")])
bm4 <- bac %>% filter(tre == "Half25C", day %in% bac_max1$day[which(bac_max1$tre=="Half25C")])

bac_max <- rbind(bm1, bm2, bm3,bm4) %>% select(tre:temp, OD, -Date) 
bac_max$nut <- factor(bac_max$nut, levels =  c("Half", "Full"))

# CV
bac_desc<- bac%>% group_by(tre,rep) %>%na.omit()%>% summarize(sd = sd(OD), meanOD= mean(OD), CV = sd/mean) %>% full_join(bac_desc) 

# mean
bac.mean_desc <- bac_desc %>% 
  group_by(tre, nut, temp) %>%
  na.omit %>%
  summarise(ig = mean(init_growth),
            CV.m=mean(CV))
bac_max_mean <- bac_max %>% group_by(nut, temp,tre) %>% na.omit() %>% summarise(ODmax = mean(OD),ODse = sd(OD))
```

#### 3.1.1 Bacteria Initial growth Linear Models

```{r bac ig model, results = "hide"}
b.ig <- lm(init_growth ~ temp , data=bac_desc);summary(b.ig)
b.ig1 <- lm(init_growth ~ nut, data=bac_desc);summary(b.ig1)
b.ig2 <- lm(init_growth ~ temp + nut, data=bac_desc);summary(b.ig2) # no significant effect from neither temp nor nut
b.ig3 <- lm(init_growth ~ temp * nut, data=bac_desc);summary(b.ig3)
```


```{r}
anova(b.ig, b.ig1,b.ig2, b.ig3)
AIC(b.ig, b.ig1,b.ig2, b.ig3)
```

#### 3.1.2 Bacteria Maximum abundance Linear Models
```{r bac OD model, results = "hide"}
b.max <- lm(OD ~ temp, bac_max);summary(b.max) # significant temperature effects
b.max1 <- lm(OD ~ nut, bac_max);summary(b.max1) # Significant nutrient effect
b.max2 <- lm(OD ~ nut+temp, bac_max);summary(b.max2) # significant nutrient and temp effect
b.max3 <- lm(OD ~ nut*temp, bac_max);summary(b.max3) # no interaction
```

```{r}
anova(b.max, b.max2,b.max1,b.max3)
AIC(b.max, b.max1,b.max2,b.max3)
```

#### 3.1.3 Bacteria Coefficient of variation (CV) Linear Models

```{r bac CV model, results = "hide"}
# coefficient of variation
b.cv <- lm(CV ~ temp, data=bac_desc);summary(b.cv) # temp not significant
b.cv1 <- lm(CV ~ nut, data=bac_desc);summary(b.cv1) # nut significant
b.cv2<- lm(CV ~ nut+temp, data=bac_desc);summary(b.cv2) # nut significant
b.cv3 <- lm(CV ~ nut*temp, data=bac_desc);summary(b.cv3) # nut significant
```


```{r}
anova(b.cv, b.cv1,b.cv2, b.cv3) # two model significantly different. Interactive model has higher R2
AIC(b.cv, b.cv1,b.cv2, b.cv3)
```


## 3.2 Tetrahymena 

```{r tetrahymena descriptor}
# new data frame
tetra_desc <- data.frame("tre" = rep(c("Full22C", "Full25C", "Half22C", "Half25C"), each = 6))
tetra_desc<-tetra_desc %>%
  mutate(temp = as.factor(rep(c(22,25), each=6, times=2)), 
         nut = factor(rep(c("Full","Half"), each=12), levels = c("Half", "Full")), 
         rep = rep(1:6, times = 4))

## initial growth rate of tetrahymena from day0 to day 1 
tetra_day0 <- tetra%>% filter(day==0) %>% select(density)
tetra_day1 <- tetra%>% filter(day==1) %>% select(density)
tetra_desc$tetra.ig <- log(tetra_day1$density)-log(tetra_day0$density)

 ## max abundance
tetra_max1 <- teb.mean %>% group_by(tre,treatNum) %>% select(tre:day,Tetrahymena) %>% summarize(max = max(Tetrahymena))
tetra_max <- teb.mean %>% filter(Tetrahymena %in% tetra_max1$max)
tetra_desc <- tetra%>% filter(tre %in% tetra_max$tre & day %in% tetra_max$day)%>% select(tre,rep,density) %>% rename(max = density)%>% left_join(tetra_desc)

# CV
tetra_desc <- tetra%>% group_by(tre,rep) %>% summarize(sd = sd(density), mean= mean(density), CV = sd/mean)%>% full_join(tetra_desc)

## time collapse
tetra_desc <- tetra %>% group_by(tre,rep) %>% filter(density == 0) %>% summarize(tc = min(day)) %>% full_join(tetra_desc)
# mean collapse time per treatment
tetra_desc %>% na.omit%>% group_by(tre) %>% summarize(mean_tc = mean(tc))

# mean of the descriptors
t.mean_desc <- tetra_desc %>%
  group_by(tre, nut, temp) %>%
  summarise(ig = mean(tetra.ig),
            max.m = mean(max), 
            CV.m = mean(CV),
            tc.m = mean(tc)) 
```

#### 3.2.1 Tetrahymena Initial growth Linear Models
 
```{r tetra ig model,results = "hide"}
# initial growth
t.ig <- lm(tetra.ig ~ temp, data=tetra_desc); summary(t.ig) #temp significant
t.ig1 <- lm(tetra.ig ~ nut, data=tetra_desc); summary(t.ig1)  # nut NOT significant
t.ig2 <- lm(tetra.ig ~ temp + nut, data=tetra_desc); summary(t.ig2) # significant temp 
t.ig3 <- lm(tetra.ig ~ temp * nut, data=tetra_desc);summary(t.ig3) # significant temp no interaction
```


```{r }
anova(t.ig,t.ig1, t.ig2,t.ig3)
AIC(t.ig,t.ig1, t.ig2,t.ig3)
```

#### 3.2.2 Tetrahymena Maximum abundance Linear Models

```{r tetra max abundance model,results = "hide"}
# max abundance
t.max <- lm(max ~ temp, data=tetra_desc); summary(t.max)
t.max1 <- lm(max ~ nut , data=tetra_desc); summary(t.max1)  # significant nutrient
t.max2 <- lm(max ~ temp + nut, data=tetra_desc); summary(t.max2) # significant nutrient, NO temp
t.max3 <- lm(max ~ temp * nut, data=tetra_desc); summary(t.max3) # significant temp and nutrient and interactive effect. Best model
```


```{r}
anova(t.max2, t.max1) # model significantly different
AIC(t.max,t.max1, t.max2,t.max3)
```
#### 3.2.3 Tetrahymena Coefficient of variation (CV) Linear Models

```{r tetra CV model,results = "hide"}
#Coefficient of varitaion
t.cv <- lm(CV ~ temp, data=tetra_desc);summary(t.cv)
t.cv1 <- lm(CV ~ nut, data=tetra_desc);summary(t.cv1)
t.cv2 <- lm(CV ~ nut+temp, data=tetra_desc);summary(t.cv2) # Temp significant. nut p = 0.08
t.cv3 <- lm(CV ~ nut*temp, data=tetra_desc);summary(t.cv3) # temp, nut and interaction significant, 0.08
```


```{r }
anova(t.cv1, t.cv2) # two model significantly different. interaction model better
AIC(t.cv, t.cv1,t.cv2, t.cv3)
```

#### 3.2.4 Tetrahymena Day of population collapse Linear Models

```{r tetra time collapse model,results = "hide"}
# time extinction/ day of collapse
tExt <- lm(tc ~ temp, data=tetra_desc); summary(tExt)  # significant Temperature effects
tExt1 <- lm(tc~  nut, data=tetra_desc);summary(tExt1)
tExt2 <- lm(tc~ temp + nut, data=tetra_desc);summary(tExt2) 
tExt3 <- lm(tc~ temp * nut, data=tetra_desc);summary(tExt3)
```


```{r }
anova(tExt1, tExt2)
AIC(tExt, tExt1, tExt2, tExt3)
```

## 3.3 Euplotes

```{r Eup descriptor, echo = FALSE}
## New data frame
eup_desc <- data.frame("tre" = rep(c("Full22C", "Full25C", "Half22C", "Half25C"), each = 6))
eup_desc<- eup_desc %>%
  mutate(temp = as.factor(rep(c(22,25), each=6, times=2)), 
         nut = factor(rep(c("Full","Half"), each=12), levels = c("Half", "Full")), 
         rep = rep(1:6, times = 4))

  ## initial growth rate of Euplotes from day0 to day 8 
eup_day0 <- eup[which(eup$day == 0), ]
eup_day8 <- eup[which(eup$day == 8), ]
eup_desc$init_growth <- (log(eup_day8$density)-log(eup_day0$density))/8

# day peak

for (i in 1:6){
for(t in 1:4) {
   eup_desc$maxdp[i+ (t-1)*6] <- max(eup$density[which(eup$treatNum == t & eup$rep==i)])
   }
} 
for (i in 1:6){
for(t in 1:4) {
   eup_desc$daypeak[i+ (t-1)*6] <- eup$day[which(eup$density== eup_desc$maxdp[i+ (t-1)*6] & eup$rep==i & eup$treatNum == t)]
   }
}

## max abundance 
eup_max1 <- teb.mean %>% group_by(tre,treatNum) %>% select(tre:day,Euplotes) %>% summarize(max = max(Euplotes))
eup_max <- teb.mean %>% filter(Euplotes %in% eup_max1$max)

eup_desc$max[1:6] <- eup$density[which(eup$day== eup_max$day[1] & eup$treatNum == 1)]
eup_desc$max[7:12] <- eup$density[which(eup$day== eup_max$day[2] & eup$treatNum == 2)]
eup_desc$max[13:18] <- eup$density[which(eup$day== eup_max$day[3] & eup$treatNum == 3)]
eup_desc$max[19:24] <- eup$density[which(eup$day== eup_max$day[4] & eup$treatNum == 4)]

# Coefficient of variation
eup_desc <- eup%>% group_by( tre,rep) %>% summarize(sd = sd(density), mean= mean(density), CV = sd/mean) %>% full_join(eup_desc)

# descriptors mean
eup.mean_desc <- eup_desc %>% 
  group_by(tre, nut, temp) %>%
  summarise(ig = mean(init_growth),
            max.m = mean(max), 
            daypeak.m = mean(daypeak),
            CV.m = mean(CV))
```

#### 3.3.1 Euplotes Initial growth Linear Models

```{r eup ig model, results = "hide"}
# Initial Growth
e.ig <- lm(init_growth ~ temp, data=eup_desc);summary(e.ig) 
e.ig1 <- lm(init_growth ~  nut, data=eup_desc);summary(e.ig1) # no nut 
e.ig2 <- lm(init_growth ~ temp + nut, data=eup_desc);summary(e.ig2) # significant temp
e.ig3 <- lm(init_growth ~ temp * nut, data=eup_desc);summary(e.ig3) # significant temp, sig nut, and interactive!
```


```{r }
anova(e.ig, e.ig1,e.ig2, e.ig3)
AIC(e.ig, e.ig1,e.ig2, e.ig3)
```
#### 3.3.2 Euplotes Maximum abundance Linear Models
```{r eup max abund model, results = "hide"}
# Maximum Abundance 
e.max <- lm(max ~ temp, data=eup_desc);summary(e.max)
e.max1 <- lm(max ~ nut, data=eup_desc);summary(e.max1)
e.max2 <- lm(max ~ nut+temp, data=eup_desc);summary(e.max2)   # significant temp and nut additive
e.max3 <- lm(max ~ nut*temp, data=eup_desc);summary(e.max3) # no interactive
```


```{r }
anova(e.max2, e.max3)
AIC(e.max, e.max1, e.max2, e.max3)
```

#### 3.3.3 Euplotes Coefficient of variation (CV) Linear Models

```{r eup CV model, results = "hide"}
# coefficient of variation
e.cv <- lm(CV ~ temp, data=eup_desc);summary(e.cv)
e.cv1 <- lm(CV ~ nut, data=eup_desc);summary(e.cv1)
e.cv2 <- lm(CV ~ nut+temp, data=eup_desc);summary(e.cv2) # nut and temp both significant
e.cv3 <- lm(CV ~ nut*temp, data=eup_desc);summary(e.cv3) # nut, temp and interactive all significant  R2 = 0.7545
```


```{r }
anova(e.cv1, e.cv2) # two model significantly different. interaction model better
AIC(e.cv,e.cv1, e.cv2,e.cv3)
```
#### 3.3.4 Euplotes Day of population peak Linear Models

```{r eup dp model, results = "hide"}
# day peaked
e.dp <- lm(daypeak ~ temp, data=eup_desc);summary(e.dp) 
e.dp1 <- lm(daypeak ~ nut, data=eup_desc);summary(e.dp1) 
e.dp2 <- lm(daypeak ~ nut+temp, data=eup_desc);summary(e.dp2) # significant temp, no nut
e.dp3 <- lm(daypeak ~ nut*temp, data=eup_desc);summary(e.dp3) # significant temp, no interavtive nor nut

```


```{r }
anova(e.dp1, e.dp2, e.dp3)
AIC(e.dp,e.dp1, e.dp2, e.dp3)
```
